Inferring generation-interval distributions from contact tracing data: \\ \emph{in prep, PRSB}

### 2019-06-25 doc/network.tex
resolving comments

The population-level egocentric kernel is found by integrating over individual variations:
\begin{equation}\label{eq:ego}
\hat{K}(\tau) = \int \hat{k}(\tau; a) \hat{f}(a) da,
\end{equation}
where $\hat{f}(a)$ represents a probability density over an aspect space, including the contact structure from an egocentric point of view (i.e., the degree distribution across network). \swp{New notation (f hat)}\jd{This feels like a mistake. We have the same population, and therefore the same distribution over aspect space, even if we are using more information to calculate the egocentric kernel.}
\cite{trapman2016inferring} used this same kernel (also assuming a constant per-pair contact rate) to study the effect of network structure on the estimate of reproductive number.
The population-level egocentric generation-interval distribution is:
\begin{equation}
\hat{g}(\tau) = \frac{\hat{K}(\tau)}{\int \hat{K}(s) ds}.
\label{eq:conditional}
\end{equation}
The population-level egocentric generation-interval distribution describes the distribution of times at which secondary infections are realized by an \emph{average} primary case; for convenience, we will often omit ``population-level''.
Finally, the link between the growth rate and the egocentric reproductive number can be obtained by applying the Euler-Lotka equation to the egocentric generation-interval distribution \citep{trapman2016inferring}:
\begin{equation}
\frac{1}{\hat{\RR}} = \int \hat{g}(\tau) \exp(-r \tau) d\tau.
\label{eq:egorR}
\end{equation}
As the egocentric distribution always has a shorter mean than the intrinsic distribution, \Rhat\ will always be smaller than $\RR$ estimated from the intrinsic distribution;
this generation-interval-based argument provides a clear biological interpretation for the result presented by \cite{trapman2016inferring}.
% Similarly, \Rhat\ will be greater than the \emph{true} reproductive number \RR\, since it does not account for depletion of susceptibles by other routes.

@@

\swp{Can you check if this subsection is clear?}
\jd{Seems clear enough. See framing below.}
Here, we show that the egocentric relationship between $r$ and $\RR$ derived by Trapman \cite{trapman2016inferring} (see the original source for detailed derivations) can be derived simply by applying the Euler-Lotka equation and using the effective (rather than the intrinsic) generation-interval distribution. 
Assume that latent and infectious periods are exponentially distributed with mean $1/\sigma$ and $1/\gamma$, respectively.
Assuming a constant per-pair contact rate of $\lambda$ for any pair, the egocentric generation-interval distribution can be written:
\begin{equation}
\hat{g}(\tau) = \frac{\sigma (\gamma + \lambda)}{\sigma - (\gamma + \lambda)} \left(e^{-(\gamma + \lambda)\tau} - e^{-\sigma \tau}\right).
\end{equation}
Substituting into \eref{egorR}, we get
\begin{equation}
\hat\RR = \left(1 + \frac{r}{\sigma}\right) \left(1 + \frac{r}{\gamma + \lambda}\right),
\end{equation}
where $r$ is the exponential growth rate.
Alternatively, the reproductive number can be expressed based on the degree distribution (mean $\mu$ and variance $\sigma^2$) of a network:
\begin{equation}
\hat\RR = \underbrace{\kappa \lambda}_{\text{average contact rate}} \times \underbrace{\frac{1}{\gamma + \lambda}}_{\text{mean effective infectious period}},
\end{equation}
where $\kappa = \sigma^2/\mu + \mu - 1$, referred to as the mean degree excess \citep{newman2003structure}, describes the expected number of susceptible individuals that an average infected individual will encounter early in an outbreak.
Combining the two equations, we get
\begin{equation}
\lambda = \frac{(\gamma + r) (\sigma + r)}{(\kappa - 1) \sigma - r},
\end{equation}
which completes the relationship between the growth rate and the egocentric reproductive number:
\begin{equation}
\hat\RR = \frac{\gamma + r}{\gamma \sigma/(\sigma + r) + r/\kappa}.
\end{equation}

>>>>

The population-level egocentric kernel is found by integrating over individual variations:
\begin{equation}\label{eq:ego}
\hat{K}(\tau) = \int \hat{k}(\tau; a) f(a) da.
\end{equation}
\cite{trapman2016inferring} used this same kernel (also assuming a constant per-pair contact rate) to study the effect of network structure on the estimate of reproductive number.
The population-level egocentric generation-interval distribution is:
\begin{equation}
\hat{g}(\tau) = \frac{\hat{K}(\tau)}{\int \hat{K}(s) ds}.
\label{eq:conditional}
\end{equation}
The population-level egocentric generation-interval distribution describes the distribution of times at which secondary infections are realized by an \emph{average} primary case; for convenience, we will often omit ``population-level''.
Finally, the link between the growth rate and the egocentric reproductive number can be obtained by applying the Euler-Lotka equation to the egocentric generation-interval distribution \citep{trapman2016inferring}:
\begin{equation}
\frac{1}{\hat{\RR}} = \int \hat{g}(\tau) \exp(-r \tau) d\tau.
\label{eq:egorR}
\end{equation}
As the egocentric distribution always has a shorter mean than the intrinsic distribution, \Rhat\ will always be smaller than $\RR$ estimated from the intrinsic distribution;
this generation-interval-based argument provides a clear biological interpretation for the result presented by \cite{trapman2016inferring}.
% Similarly, \Rhat\ will be greater than the \emph{true} reproductive number \RR\, since it does not account for depletion of susceptibles by other routes.

@@

Here, we show that the egocentric relationship between $r$ and $\RR$ derived by Trapman \cite{trapman2016inferring} (see the original source for detailed derivations) can be derived simply by applying the Euler-Lotka equation and using the effective (rather than the intrinsic) generation-interval distribution. 
Assume that latent and infectious periods are exponentially distributed with mean $1/\sigma$ and $1/\gamma$, respectively.
Assuming a constant per-pair contact rate of $\lambda$ for any pair, the egocentric generation-interval distribution can be written:
\begin{equation}
\hat{g}(\tau) = \frac{\sigma (\gamma + \lambda)}{\sigma - (\gamma + \lambda)} \left(e^{-(\gamma + \lambda)\tau} - e^{-\sigma \tau}\right).
\end{equation}
Substituting into \eref{egorR}, we get
\begin{equation}
\hat\RR = \left(1 + \frac{r}{\sigma}\right) \left(1 + \frac{r}{\gamma + \lambda}\right),
\end{equation}
where $r$ is the exponential growth rate.
Alternatively, the reproductive number can be expressed based on the degree distribution (mean $\mu$ and variance $\sigma^2$) of a network:
\begin{equation}
\hat\RR = \underbrace{\kappa \lambda}_{\text{average contact rate}} \times \underbrace{\frac{1}{\gamma + \lambda}}_{\text{mean effective infectious period}},
\end{equation}
where $\kappa = \sigma^2/\mu + \mu - 1$, referred to as the mean degree excess \citep{newman2003structure}, describes the expected number of susceptible individuals that an average infected individual will encounter early in an outbreak.
Combining the two equations, we get
\begin{equation}
\lambda = \frac{(\gamma + r) (\sigma + r)}{(\kappa - 1) \sigma - r},
\end{equation}
which completes the relationship between the growth rate and the egocentric reproductive number:
\begin{equation}
\hat\RR = \frac{\gamma + r}{\gamma \sigma/(\sigma + r) + r/\kappa}.
\end{equation}
